---
title: "ESM 244 - Assignment 2"
author: "Alexander Stejskal"
date: "February 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1. Data wrangling and visualization - cetaceans in capitvity

## Part 1 - Add packages and import data

Add packages
```{r}
library(readr)
library(tidyverse)
library(ggplot2)

```


Import dataset "captive_cetacean.csv"
```{r}
cetacean <- read_csv("~/winter 2019/esm 244 - adv data/assignment 2/244_hw2/captive_cetacean.csv")
```


## Part 2 - Data wrangling and Exploration

Explore data
```{r}
head(cetacean)
```

View the data types
```{r}
str(cetacean)
```


Explore the results within the "species" column
```{r}
count(cetacean, acquisition)
```


Explore the results within the "status" column
```{r}
count(cetacean, status)
```

Filter 'status' column so only "Alive" and "Dead" are shown
```{r}
df_status <- filter(cetacean, status == "Alive"| status == "Died")
```


Filter 'acquisiton' column in 'df_status' so only results for 'Born', 'Capture', and 'Rescue' are shown
```{r}
df_acq_status <- filter(df_status, acquisition == 'Born'| acquisition == 'Capture' | acquisition == 'Rescue')
```



## Part 3 - Visualize data

Create histogram on the categorical variable 'acquisition' to display the number of animals that are alive or have died based on their method of acquisition 
```{r}
acq_status_hist <- ggplot(df_acq_status, aes(acquisition)) +
  geom_bar(aes(fill = status), width = 0.5) +
  ggtitle("Cetacean Mortality by Acquisition Method\n") +
  xlab("Acquisition Method") +
  ylab("Count") +
  labs(fill = "Status") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(color = "black"), panel.background = element_rect(fill = "white")) +
  scale_y_continuous(expand = c(0,0))+
  theme(plot.title = element_text(color = "black", size = 15, face = "bold", hjust = 0.5)) +
  theme(axis.title = element_text(size = 12))+
  theme(axis.text = element_text(size = 10, color = "black")) +
  theme(legend.title = element_text(color = "black"))
  

acq_status_hist

```



# Task 2 - Parameter Estimation - Wild Fish Catch


## Part 1 - Add packages and import data

Import dataset "fish_catch.csv"
```{r}
fish_catch <- read_csv("~/winter 2019/esm 244 - adv data/assignment 2/244_hw2_22/244_hw2_2_17_2019/fish_catch.csv")
```


Add packages
```{r}
library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car)
library(latex2exp)
```


## Part 2 - Data Wrangling and Exploration

Rename columns in 'fish_catch'
'TFP' = 'Total Fish Production'
```{r}
colnames(fish_catch) <- c("year", "wild_catch", "farmed_fish", "TFP")
```


Remove the first and last 4 rows of "fish_catch"
```{r}
df_fish <- fish_catch[-c(1:4,68:71),]
```


Add column with index number for each year (ex. 1950 = 0, 1951 = 1, etc.)
```{r}
df_fish$index_year <- 0:62
```

Change columns to numeric
```{r}
df_fish[,2:4] <- sapply(df_fish[,2:4], as.numeric)
```

Remove unnecessary columns
```{r}
df_fish$year <- NULL
df_fish$farmed_fish <- NULL
df_fish$TFP <- NULL
```


### A)

Create a graph of wild catch over time
```{r}
ggplot(df_fish, aes(x = index_year, y = wild_catch)) + 
  geom_point() +
  theme_bw() +
  scale_y_continuous(breaks = c(20,40,60,80,100)) +
  annotate(geom = 'text', x = 50,y = 50, label = TeX("N(t) = $\\frac{90}{\\1 + 4.3e^{\\-0.05225 t}}", output = 'character'),parse=TRUE)

```

The data appears to follow a logistic growth trend. 

Logistic growth equation: 

N(t) = A/(1+Be^-rt)

# Estimate for K/A: ~ 90
# Estimate for N0: ~ 17
# Estimate for B: ~ 4.3

BEstimate <- (90 - 17)/17 # BEstimate = 4.3



Create a subset of the data that you think is just in the 'exponential growth phase.' Take the natural log of the count and create a basic scatterplot of time v ln(wild_catch)

```{r}
graph1 <- plot(df_fish$index_year[0:10],log(df_fish$wild_catch[0:10]))
```



Then find the slope of that line to get 'r' (growth rate constant)
```{r}
r_est <- lm(log(df_fish$wild_catch[0:10]) ~ df_fish$index_year[0:10]) # R ~ 0.05225

#r_est
```


Estimated logistic growth equation:

N(t) = 90/(1+4.3e^(-0.05225)t)



```{r}
plot(TeX("N(t) = $\\frac{90}{\\1 + 4.3e^{\\-0.05225 t}}"))
```































